---
title: "ESM204 HW4 Andrew, Qusai, Tor"
author: "Andrew Salvador, Qusai Bhaijeewala, & Tor Fibiger"
date: "5/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```


```{r}
# Attach necessary packages
library(janitor)
library(tidyverse)
library(broom)
```

Build the functions, set variables t & T, everything else is given
temp = temperature at year t
K_t = fraction of economic activity retained at year t, if temperature is T(t)
C_t = economic activity ("consumption") at time t
U_c = society's utility from consumption
discount rate r = delta + ng
```{r}
t = 100
T = 4.4
B = 0.05
g = 0.01
n = 0.50
delta = 0.005
r = delta + n*g
#temp = tauf

#FUNCTION 1 (temperature at year 't', temperature 'T' at year 100)
tauf = function(t,T)
{
  out = min(t*T/100, T)
  return(out)
}


#FUNCTION 2 (fraction of economic activity)
Keep = function(temp)
{
  out = exp(-B*temp^2)
  return(out)
}


#FUNCTION 3 (economic activity "consumption", grows at rate g, reduced by K)
Consumption = function(t,temp)
{
  out = Keep(temp)*exp(g*t)
  return(out)
}


#FUNCTION 4 (society's utility from consumption)
Utility = function(C)
{
  out = ((C)^(1-n))/(1-n)
  return(out)
}

```

--------------------------------------------------------------------------

### QUESTION 1: PLOTS
(a) Plot temperature over time for no climate change (T = 0), modest climate change (T = 2), and with extreme climate change (T = 8).


```{r}
# Temperature without climate change (T = 0) (wocc = without climate change)

t = 200
T = 0
B = 0.05
g = 0.01
n = 0.50
delta = 0.005

DF_T_wocc = data.frame(t=seq(0,200)) %>% 
  rowwise() %>% 
  mutate(temperature = tauf(t,T)) %>% 
  mutate(K = Keep(temperature)) %>% 
  mutate(C = Consumption(t,temperature)) %>% 
  mutate(U = Utility(C)) %>% 
  mutate(PVU = U/(1+r)^t) %>% 
  select(t, temperature)

plot(DF_T_wocc,
     main = "Temperature without climate change (T = 0)",
     xlab = "Year",
     ylab = "Temperature")


```


```{r}
# Temperature with modest climate change (T = 2) (mcc = modest climate change)
t = 200
T = 2
B = 0.05
g = 0.01
n = 0.50
delta = 0.005

DF_T_mcc = data.frame(t=seq(0,200)) %>% 
  rowwise() %>% 
  mutate(temperature = tauf(t,T)) %>% 
  mutate(K = Keep(temperature)) %>% 
  mutate(C = Consumption(t,temperature)) %>% 
  mutate(U = Utility(C)) %>% 
  mutate(PVU = U/(1+r)^t) %>% 
  select(t, temperature)

plot(DF_T_mcc,
     main = "Temperature with moderate climate change (T = 2)",
     xlab = "Year",
     ylab = "Temperature")
```

```{r}
# Temperature with extreme climate change (T = 8) (ecc = extreme climate change)
t = 200
T = 8
B = 0.05
g = 0.01
n = 0.50
delta = 0.005

DF_T_ecc = data.frame(t=seq(0,200)) %>% 
  rowwise() %>% 
  mutate(temperature = tauf(t,T)) %>% 
  mutate(K = Keep(temperature)) %>% 
  mutate(C = Consumption(t,temperature)) %>% 
  mutate(U = Utility(C)) %>% 
  mutate(PVU = U/(1+r)^t) %>% 
  select(t, temperature)

plot(DF_T_ecc,
     main = "Temperature with extreme climate change (T = 8)",
     xlab = "Year",
     ylab = "Temperature")
```

(b) Plot consumption over time for no climate change, modest climate change, and extreme climate change.

```{r}
# Consumption without climate change (C_wocc = consumption without climate change)

t = 200
T = 0
B = 0.05
g = 0.01
n = 0.50
delta = 0.005

DF_C_wocc = data.frame(t=seq(0,200)) %>% 
  rowwise() %>% 
  mutate(temperature = tauf(t,T)) %>% 
  mutate(K = Keep(temperature)) %>% 
  mutate(C = Consumption(t,temperature)) %>% 
  mutate(U = Utility(C)) %>% 
  mutate(PVU = U/(1+r)^t) %>% 
  select(t,C)

plot(DF_C_wocc,
     main = "Consumption without climate change (T = 0)",
     xlab = "Year",
     ylab = "Consumption")
```

```{r}
# Consumption modest climate change (T = 2) (C_mcc = consumption modest climate change)
t = 200
T = 2
B = 0.05
g = 0.01
n = 0.50
delta = 0.005

DF_C_mcc = data.frame(t=seq(0,200)) %>% 
  rowwise() %>% 
  mutate(temperature = tauf(t,T)) %>% 
  mutate(K = Keep(temperature)) %>% 
  mutate(C = Consumption(t,temperature)) %>% 
  mutate(U = Utility(C)) %>% 
  mutate(PVU = U/(1+r)^t) %>% 
  select(t,C)

plot(DF_C_mcc,
     main = "Consumption with moderate climate change (T = 2)",
     xlab = "Year",
     ylab = "Consumption")
```

```{r}
# Consumption extreme climate change (T = 8) (C_ecc = consumption extreme climate change)
t = 200
T = 8
B = 0.05
g = 0.01
n = 0.50
delta = 0.005

DF_C_ecc = data.frame(t=seq(0,200)) %>% 
  rowwise() %>% 
  mutate(temperature = tauf(t,T)) %>% 
  mutate(K = Keep(temperature)) %>% 
  mutate(C = Consumption(t,temperature)) %>% 
  mutate(U = Utility(C)) %>% 
  mutate(PVU = U/(1+r)^t) %>% 
  select(t,C)

plot(DF_C_ecc,
     main = "Consumption with extreme climate change (T = 8)",
     xlab = "Year",
     ylab = "Consumption")
```

(c) Plot undiscoutned utility over time for no climate change, modest climate change, and extreme climate change.

```{r}
# Undiscounted utility without climate change (T = 0) 
t = 200
T = 0
B = 0.05
g = 0.01
n = 0.50
delta = 0.005

DF_U_wocc = data.frame(t=seq(0,200)) %>% 
  rowwise() %>% 
  mutate(temperature = tauf(t,T)) %>% 
  mutate(K = Keep(temperature)) %>% 
  mutate(C = Consumption(t,temperature)) %>% 
  mutate(U = Utility(C)) %>% 
  mutate(PVU = U/(1+r)^t) %>% 
  select(t,U)

plot(DF_U_wocc,
     main = "Undiscounted utility without climate change (T = 0)",
     xlab = "Year",
     ylab = "Undiscounted utility")
```

```{r}
# Plot utility with modest climate change (T = 2) 
t = 200
T = 2
B = 0.05
g = 0.01
n = 0.50
delta = 0.005

DF_U_mcc= data.frame(t=seq(0,200)) %>% 
  rowwise() %>% 
  mutate(temperature = tauf(t,T)) %>% 
  mutate(K = Keep(temperature)) %>% 
  mutate(C = Consumption(t,temperature)) %>% 
  mutate(U = Utility(C)) %>% 
  mutate(PVU = U/(1+r)^t) %>% 
  select(t,U)

plot(DF_U_mcc,
     main = "Undiscounted utility with moderate climate change (T = 2)",
     xlab = "Year",
     ylab = "Undiscounted utility")
```

```{r}
# Plot utility with extreme climate change (T = 8) 
t = 200
T = 8
B = 0.05
g = 0.01
n = 0.50
delta = 0.005

DF_U_ecc = data.frame(t=seq(0,200)) %>% 
  rowwise() %>% 
  mutate(temperature = tauf(t,T)) %>% 
  mutate(K = Keep(temperature)) %>% 
  mutate(C = Consumption(t,temperature)) %>% 
  mutate(U = Utility(C)) %>% 
  mutate(PVU = U/(1+r)^t) %>% 
  select(t,U)

plot(DF_U_ecc,
     main = "Undiscounted utility with extreme climate change (T = 8)",
     xlab = "Year",
     ylab = "Undiscounted utility")
```

----------------------------------------------------------------------------------------

### QUESTION 2: ANALYSIS

#### (a) Suppose T = 4.4. What is the present value (i.e. discounted) utility over the next 200 years with climate change? What is the present value utility without climate change? What is the percentage loss in present value utility from climate change?

```{r}
#This dataframe shows values for temperature, economic activity retained (K), consumption (C), utility (U), and present value utility (PVU) from years 0 to 200. It also gives the NPV which is the sum of all the PVUs over those years. Use this to compare PVU with and without climate change in question 2(a,b). 

# With climate change (wcc)
t = 200
T = 4.4
B = 0.05
g = 0.01
n = 0.50
delta = 0.005

wcc_DF = data.frame(t=seq(0,200)) %>% 
  rowwise() %>% 
  mutate(temperature = tauf(t,T)) %>% 
  mutate(K = Keep(temperature)) %>% 
  mutate(C = Consumption(t,temperature)) %>% 
  mutate(U = Utility(C)) %>% 
  mutate(PVU = U/(1+r)^t) 

  wcc_NPV = sum(wcc_DF$PVU)
  
# Net present value utility = 198.661219717316
```
Net present value utility with climate change = 198.661219717316


```{r}
# Without climate change (wocc)
t = 200
T = 0
B = 0.05
g = 0.01
n = 0.50
delta = 0.005

wocc_DF = data.frame(t=seq(0,200)) %>% 
  rowwise() %>% 
  mutate(temperature = tauf(t,T)) %>% 
  mutate(K = Keep(temperature)) %>% 
  mutate(C = Consumption(t,temperature)) %>% 
  mutate(U = Utility(C)) %>% 
  mutate(PVU = U/(1+r)^t) 

  wocc_NPV = sum(wocc_DF$PVU)
  
# Net present value utility = 255.273441757037
```
Net present value utility without climate change = 255.273441757037

```{r}
# To calculate L:
# 255.273441757037 - 198.661219717316 = 56.61222
# 56.61222/255.273441757037 = 0.2217709 = 22.18%
# *L = 22.18*

L = (wocc_NPV - wcc_NPV)/wocc_NPV*100
```
L = 22.18


#### (b) Now, show how sensitive your calculation of L is to the following parameters: T, g, n, and B. To do so, calculate the % change in L that arises from a 10% increase in each of these parameters.

**Sensitivity to T** 

*With climate change sensitivity to T (10% increase of 4.4, new T = 4.84)* 
```{r}
# With climate change sensitivity to T (10% increase of 4.4, new T = 4.84) (wcc_NPV_sT)
t = 200
T = 4.84
B = 0.05
g = 0.01
n = 0.50
delta = 0.005

wcc_DF_sT = data.frame(t=seq(0,200)) %>% 
  rowwise() %>% 
  mutate(temperature = tauf(t,T)) %>% 
  mutate(K = Keep(temperature)) %>% 
  mutate(C = Consumption(t,temperature)) %>% 
  mutate(U = Utility(C)) %>% 
  mutate(PVU = U/(1+r)^t) 

  wcc_NPV_sT = sum(wcc_DF_sT$PVU)
  
# Net present value utility = 189.421099192601
```
New net present value utility with climate change = 189.421099192601

Without climate change sensitivity to T (10% increase of 0, new T = 0) 
```{r}
# Without climate change sensitivity to T (10% increase of 0, new T = 0) (wocc_NPV_sT)
t = 200
T = 0
B = 0.05
g = 0.01
n = 0.50
delta = 0.005

wocc_DF_sT = data.frame(t=seq(0,200)) %>% 
  rowwise() %>% 
  mutate(temperature = tauf(t,T)) %>% 
  mutate(K = Keep(temperature)) %>% 
  mutate(C = Consumption(t,temperature)) %>% 
  mutate(U = Utility(C)) %>% 
  mutate(PVU = U/(1+r)^t) 

  wocc_NPV_sT = sum(wocc_DF_sT$PVU)
  
# Net present value utility = 255.273441757037
```
New net present value utility without climate change = 255.273441757037


```{r}
# To calculate L
# 255.273441757037 - 189.421099192601 = 65.85234
# 65.85234/255.273441757037 = 0.2579678 = 25.80
# *New L = 25.80%*
# **A 10% increase in T caused L to increased from 22.18% to 25.80%**

L_T = (wocc_NPV_sT - wcc_NPV_sT)/wocc_NPV_sT*100
```
*New L = 25.80%*

~~~~~~~
A 10% increase in T caused L to increased from 22.18% to 25.80%, a 16.32% increase.
~~~~~~~

**Sensitivity to g**

*With climate change sensitivity to g (10% increase to 0.01, new g = 0.011)* 
```{r}
# With climate change sensitivity to g (10% increase to 0.01, new g = 0.011) (wcc_NPV_sg)
t = 200
T = 4.4
B = 0.05
g = 0.011
n = 0.50
delta = 0.005

wcc_DF_sg = data.frame(t=seq(0,200)) %>% 
  rowwise() %>% 
  mutate(temperature = tauf(t,T)) %>% 
  mutate(K = Keep(temperature)) %>% 
  mutate(C = Consumption(t,temperature)) %>% 
  mutate(U = Utility(C)) %>% 
  mutate(PVU = U/(1+r)^t) 

  wcc_NPV_sg = sum(wcc_DF_sg$PVU)
  
# Net present value utility = 206.137459318153
```
New net present value utility with climate change = 206.137459318153

```{r}
# Without climate change sensitivity to g (10% increase to 0.01, new g = 0.011) (wocc_NPV_sg)
t = 200
T = 0
B = 0.05
g = 0.011
n = 0.50
delta = 0.005

wocc_DF_sg = data.frame(t=seq(0,200)) %>% 
  rowwise() %>% 
  mutate(temperature = tauf(t,T)) %>% 
  mutate(K = Keep(temperature)) %>% 
  mutate(C = Consumption(t,temperature)) %>% 
  mutate(U = Utility(C)) %>% 
  mutate(PVU = U/(1+r)^t) 

  wocc_NPV_sg = sum(wocc_DF_sg$PVU)
  
# Net present value utility = 266.277404501181
```
New net present value utility without climate change = 266.277404501181


```{r}
# To calculate L
# 266.277404501181 - 206.137459318153 = 60.13995
# 60.13995/266.277404501181 = 0.2258545 = 22.59%
# *New L = 22.59%*
# **A 10% increase in g caused L to increase minimally from  22.18% to 22.59%**

L_g = (wocc_NPV_sg - wcc_NPV_sg)/wocc_NPV_sg*100
```
*New L = 22.59%*

~~~~~~~
A 10% increase in g caused L to increase minimally from 22.18% to 22.59%, a 1.85% increase.
~~~~~~~



**Sensitivity to n**

*With climate change sensitivity to n (10% increase to 0.5, new n = 0.55)*
```{r}
# With climate change sensitivity to n (10% increase to 0.5, new n = 0.55) (wcc_NPV_sn)
t = 200
T = 4.4
B = 0.05
g = 0.01
n = 0.55
delta = 0.005

wcc_DF_sn = data.frame(t=seq(0,200)) %>% 
  rowwise() %>% 
  mutate(temperature = tauf(t,T)) %>% 
  mutate(K = Keep(temperature)) %>% 
  mutate(C = Consumption(t,temperature)) %>% 
  mutate(U = Utility(C)) %>% 
  mutate(PVU = U/(1+r)^t) 

  wcc_NPV_sn = sum(wcc_DF_sn$PVU)
  
# Net present value utility = 217.796047447958
```
New net present value utility with climate change = 217.796047447958

Without climate change sensitivity to n (10% increase to 0.5, new n = 0.55)
```{r}
# Without climate change sensitivity to n (10% increase to 0.5, new n = 0.55) (wocc_NPV_sn)
t = 200
T = 0
B = 0.05
g = 0.01
n = 0.55
delta = 0.005

wocc_DF_sn = data.frame(t=seq(0,200)) %>% 
  rowwise() %>% 
  mutate(temperature = tauf(t,T)) %>% 
  mutate(K = Keep(temperature)) %>% 
  mutate(C = Consumption(t,temperature)) %>% 
  mutate(U = Utility(C)) %>% 
  mutate(PVU = U/(1+r)^t) 

  wocc_NPV_sn = sum(wocc_DF_sn$PVU)
  
# Net present value utility = 272.133835059214
```
New net present value utility without climate change = 272.133835059214


```{r}
# To calculate L
# 272.133835059214 - 217.796047447958 = 54.33779
# 54.33779/272.133835059214 = 0.199673 = 19.97%
# *New L = 19.97%*
# **A 10% increase in n caused L to decrease from 22.18% to 19.97**

L_n = (wocc_NPV_sn - wcc_NPV_sn)/wocc_NPV_sn*100
```
*New L = 19.97%*

~~~~~~~
A 10% increase in n caused L to decrease from 22.18% to 19.97
~~~~~~~


**Sensitivity to B**

*With climate change sensitivity to B (10% increase to 0.05, new B = 0.055)*
```{r}
# With climate change sensitivity to B (10% increase to 0.05, new B = 0.055) (wcc_NPV_sB)
t = 200
T = 4.4
B = 0.055
g = 0.01
n = 0.50
delta = 0.005

wcc_DF_sB = data.frame(t=seq(0,200)) %>% 
  rowwise() %>% 
  mutate(temperature = tauf(t,T)) %>% 
  mutate(K = Keep(temperature)) %>% 
  mutate(C = Consumption(t,temperature)) %>% 
  mutate(U = Utility(C)) %>% 
  mutate(PVU = U/(1+r)^t) 

  wcc_NPV_sB = sum(wcc_DF_sB$PVU)
  
# Net present value utility = 194.164520791888
```
New net present value utility with climate change = 194.164520791888

Without climate change sensitivity to B (10% increase to 0.05, new B = 0.055)
```{r}
# Without climate change sensitivity to B (10% increase to 0.05, new B = 0.055) (wocc_NPV_sB)
t = 200
T = 0
B = 0.055
g = 0.01
n = 0.50
delta = 0.005

wocc_DF_sB = data.frame(t=seq(0,200)) %>% 
  rowwise() %>% 
  mutate(temperature = tauf(t,T)) %>% 
  mutate(K = Keep(temperature)) %>% 
  mutate(C = Consumption(t,temperature)) %>% 
  mutate(U = Utility(C)) %>% 
  mutate(PVU = U/(1+r)^t) 

  wocc_NPV_sB = sum(wocc_DF_sB$PVU)
  
# Net present value utility = 255.273441757037
```
New net present value utility without = 255.273441757037


```{r}
# To calculate L
# 255.273441757037 - 194.164520791888 = 61.10892
# 61.10892/255.273441757037 = 0.2393861 = 23.94%
# *New L = 23.94%*
# **A 10% increase in B caused L to increase from 22.18% to 23.94%**

L_B = (wocc_NPV_sB - wcc_NPV_sB)/wocc_NPV_sB*100
```
*New L = 23.94%*

~~~~~~~
A 10% increase in B caused L to increase from 22.18% to 23.94%
~~~~~~~


#### (c) Back to the original parameters, suppose we could completely prevent climate change from ocurring (so T = 0 instead of T = 4.4) but doing so would require giving up a fraction omega of consumption every year for the next 200 years. What is the maximum value of omega society would be willing to endure every year to completely prevent climate change? Call this omegastar. 

With climate change (T = 4.4), NPV utility = 198.66

Without climate change (T = 0), NPV utility = 255.27

```{r}
# Consumption without climate change (T = 0), giving up fraction omega each year
# At omega 0.53, total consumption over 200 years is higher, but NPV utility and undiscounted utility is lower
# at omega 0.39, NPV, U, and C are all higher than with climate change. 

t = 200
T = 0
B = 0.05
g = 0.01
n = 0.50
delta = 0.005
omega = 0.38

Q2cwocc_DF_omega = data.frame(t=seq(0,200)) %>% 
  rowwise() %>% 
  mutate(temperature = tauf(t,T)) %>% 
  mutate(K = Keep(temperature)) %>% 
  mutate(C = Consumption(t,temperature)-(omega*Consumption(t,temperature))) %>% 
  mutate(U = Utility(C)) %>% 
  mutate(PVU = U/(1+r)^t) 

Q2cwocc_NPV_omega = sum(Q2cwocc_DF_omega$PVU)
 

```

```{r}
# Consumption with climate change (T = 4.4)
t = 200
T = 4.4
B = 0.05
g = 0.01
n = 0.50
delta = 0.005

Q2cwcc_DF = data.frame(t=seq(0,200)) %>% 
  rowwise() %>% 
  mutate(temperature = tauf(t,T)) %>% 
  mutate(K = Keep(temperature)) %>% 
  mutate(C = Consumption(t,temperature)) %>% 
  mutate(U = Utility(C)) %>% 
  mutate(PVU = U/(1+r)^t) 

Q2cwcc_NPV = sum(Q2cwcc_DF$PVU)

# Just to look at a plot of consumption with climate change T = 4.4
t = 200
T = 4.4
B = 0.05
g = 0.01
n = 0.50
delta = 0.005

DF_C_wcc = data.frame(t=seq(0,200)) %>% 
  rowwise() %>% 
  mutate(temperature = tauf(t,T)) %>% 
  mutate(K = Keep(temperature)) %>% 
  mutate(consumption = Consumption(t,temperature)) %>% 
  mutate(U = Utility(consumption)) %>% 
  mutate(PVU = U/(1+r)^t) %>% 
  select(t,consumption)

```
 ~~~~~~~
 Omega* = 0.38
 ~~~~~~~
 

#### (d) Suppose we are uncertain about T, but it has the following probability distribution: T = 2 (probability 0.2), T = 4 (probability 0.5), T = 6 (probability 0.3). Calculate omegastar under uncertainty over T. 

```{r}
# Call omegastar 0.39 when T = 4.4
# T = 2 (20% probability)
t = 200
T = 2
B = 0.05
g = 0.01
n = 0.50
delta = 0.005

T2_DF = data.frame(t=seq(0,200)) %>% 
  rowwise() %>% 
  mutate(temperature = tauf(t,T)) %>% 
  mutate(K = Keep(temperature)) %>% 
  mutate(C = Consumption(t,temperature)) %>% 
  mutate(U = Utility(C)) %>% 
  mutate(PVU = U/(1+r)^t) 

NPV_T2 = sum(T2_DF$PVU)

# T = 4 (50% probability)
t = 200
T = 4
B = 0.05
g = 0.01
n = 0.50
delta = 0.005

T4_DF = data.frame(t=seq(0,200)) %>% 
  rowwise() %>% 
  mutate(temperature = tauf(t,T)) %>% 
  mutate(K = Keep(temperature)) %>% 
  mutate(C = Consumption(t,temperature)) %>% 
  mutate(U = Utility(C)) %>% 
  mutate(PVU = U/(1+r)^t) 

NPV_T4 = sum(T4_DF$PVU)

# T = 6 (30% probability)
t = 200
T = 6
B = 0.05
g = 0.01
n = 0.50
delta = 0.005

T6_DF = data.frame(t=seq(0,200)) %>% 
  rowwise() %>% 
  mutate(temperature = tauf(t,T)) %>% 
  mutate(K = Keep(temperature)) %>% 
  mutate(C = Consumption(t,temperature)) %>% 
  mutate(U = Utility(C)) %>% 
  mutate(PVU = U/(1+r)^t) 

NPV_T6 = sum(T6_DF$PVU)
```


```{r}
# If we look at expected value NPV, then:
expected_value = NPV_T2*0.2 + NPV_T4*0.5 + NPV_T6*0.3

#expected value = 201.3446

# omegastar must result in NPV w/no climate change > than expected value
t = 200
T = 0
B = 0.05
g = 0.01
n = 0.50
delta = 0.005
omega = 0.37
# at omega = 0.37, NPV without climate change > expected value 

Q2cwocc_DF_omega = data.frame(t=seq(0,200)) %>% 
  rowwise() %>% 
  mutate(temperature = tauf(t,T)) %>% 
  mutate(K = Keep(temperature)) %>% 
  mutate(C = Consumption(t,temperature)-(omega*Consumption(t,temperature))) %>% 
  mutate(U = Utility(C)) %>% 
  mutate(PVU = U/(1+r)^t) 

Q2dwocc_NPV_omega = sum(Q2cwocc_DF_omega$PVU)

```
~~~~~~~
Omega* = 0.37 under uncertainty over T.
~~~~~~~

